Product plots

### 2011-06-23 reviews/response.txt

We are grateful for the helpful feedback provided by the reviewers. We have added a related work section, including many of the missing references we missed. We have also weakened our claims about the superiority of rectangles, instead providing a more pragmatic justification

Other changes:

* Rather than eliminating the section describing the R implementation altogether, we have expanded it slightly to give a taste of how it might be used.

* Where appropriate, figures labelled with distributional decomposition.

### 2011-06-23 prodplots-infovis.tex
More minor fixes

\abstract{

We propose a new framework for visualising tables of counts, proportions and probabilities. We call our framework product plots, alluding to the computation of area as a product of height and width, and the statistical concept of generating a joint distribution from the product of conditional and marginal distributions. The framework, with extensions, is sufficient to encompass over 20 visualisations previously described in fields of statistical graphics and infovis, including bar charts, mosaic plots, treemaps, equal area plots and fluctuation diagrams.

}

@@

  \item {\bf spine}s: width is proportional to value, height occupies full range. Spines are space filling and can be arranged horizontally (``hspine''), vertically (``vspine''), or can automatically pick their orientation (``spine'') by splitting the largest dimension. The name spine is evocative of books sitting on a library shelf \citep{hummel:1996}. Bars and spines are indistinguishable when the underlying data is uniformly distributed.

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.9 \linewidth]{part-fluct}
    \includegraphics[width=0.9 \linewidth]{part-fluct-cond}
  \caption{(Left) A fluctuation diagram showing distribution of age (in decades) and marital status. (Right) Equal bin-size plot showing health status and survey year. Thee three empty columns show that health status was not recorded for three years.}
  \label{fig:fluct}
\end{figure}

@@

To construct plots of higher-dimensional data sets, we need a way to decompose them into 1d and 2d components. Some statistical vocabulary is useful: one way of describing the input data is as a {\bf probability mass function}, or {\sc pmf}. A {\sc pmf} is a function with n inputs, each indexing one dimension with an integer, and gives the probability of each combination of inputs. A {\sc pmf} has two restrictions: every value must be greater than or equal to zero, and all values must sum to one.

@@

  \item {\bf Infoslices} \citep{andrews:1998} (nd): n vbars. But they only use half of the polar plane, and are specialised for highly nested data.

@@

  \caption{(Top) Area graphics in Cartesian coordinates. (Mid) Area graphics in polar coordinates. From left to right: hbar + hbar, vspline + hbar, hspine + hspine, hspine + vspine. (Bottom) Product graphics rotated 90 degrees before being converted to polar coordinates: vbar + vbar, hspline + vbar, vspine + vspine, vspine + hspine. (\key{male} male, \key{female} female)}

>>>>

\abstract{We propose a new framework for visualising tables of counts, proportions and probabilities. We call our framework product plots, alluding to the computation of area as a product of height and width, and the statistical concept of generating a joint distribution from the product of conditional and marginal distributions. The framework, with extensions, is sufficient to encompass over 20 visualisations previously described in fields of statistical graphics and infovis, including bar charts, mosaic plots, treemaps, equal area plots and fluctuation diagrams.}

@@

  \item {\bf spine}s: width is proportional to value, height occupies full range. Spines are space filling and can be arranged horizontally (``hspine''), vertically (``vspine''), or can automatically pick their orientation (``spine'') by splitting the largest dimension. The name spine is evocative of books sitting on a library shelf \citep{hummel:1996}. Bars and spines are indistinguishable when the underlying data is evenly distributed across the categories.

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.9 \linewidth]{part-fluct}
    \includegraphics[width=0.9 \linewidth]{part-fluct-cond}
  \caption{(Top) A fluctuation diagram showing distribution of age (in decades) and marital status. (Bottom) Equal bin-size plot showing health status and survey year. Thee three empty columns show that health status was not recorded for three years.}
  \label{fig:fluct}
\end{figure}

@@

To construct plots of higher-dimensional data sets, we need a way to decompose them into 1d and 2d components. Some statistical vocabulary is useful: one way of describing the input data is as a {\bf probability mass function}, or {\sc pmf}. A {\sc pmf} is a function with $n$ inputs, each indexing one dimension with an integer, and gives the probability of each combination of inputs. A {\sc pmf} has two restrictions: every value must be greater than or equal to zero, and all values must sum to one.

@@

  \item {\bf Infoslices} \citep{andrews:1998} (nd): $n$ vbars. But they only use half of the polar plane, and are specialised for highly nested data.

@@

  \caption{(Top) Area graphics in Cartesian coordinates. (Mid) Area graphics in polar coordinates. (Bottom) Product graphics with alternate mapping of $x$ and $y$ to $r$ and $\theta$. From left to right: hbar + hbar, vspline + hbar, hspine + hspine, hspine + vspine. (\key{male} male, \key{female} female)}
